---
title: "Model Convergence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model Convergence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(MplusReadR)
library(MplusAutomation)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(knitr)
```

## Checking Models Convergence
You can check whether each model in a list of Mplus models converged using `mplus_converge()`. 
This checks whether the model contains confidence intervals, as non-converged models will not. 


```{r, message = FALSE, results = 'hide'}
# Load the data
bivariate_models <- readModels('bivariate_models', recursive = TRUE)
```

```{r}
# Check model convergence
converged <- mplus_converge(bivariate_models)
```

```{r, echo = FALSE}
# Print table 
kable(converged)
```




## Removing Non-converged Models 

Non-converged models in a list of Mplus output can be removed using `mplus_remove_converge()`.

```{r}
new_models <- mplus_remove_converge(bivariate_models)

```


* Number of models **before** removal of non-converged models: `r length(bivariate_models)`  
* Number of models **after** removal of non-converged models: `r length(new_models)`    

Non-converged models are removed automatically when model output is compiled using `mplus_compile()`. 
If there were any non-converged models, a warning will be displayed, as shown below.

```{r, results = 'hide'}
mplus_compile(bivariate_models)
```

You can prevent automatic removal of non-converged models using `converged = FALSE`. However, if the output contains any non-converged models,
not removing them will prevent the compilation of the model output. 
